Modelling the Co-evolution of Trade and Culture in Past Societies

### 2015-07-01 article.tex
work again on the writting of section 4.1

We observe that the lower the mutation rate, the closer to a line the result is. This line corresponds to the ``power law" distribution explained in section~\ref{sec:culturalTrans}, and is typical of the result obtained under the ``neutral hypothesis". In order to verify if the distribution is in fact a power-law, we follow the method proposed by~\cite{clauset2009powerlawdistributionsinempiricaldat} and the R implementation proposed by~\cite{gillespie_fitting_2015}. Briefly, two values are returned: a) the estimation of the $\alpha$ parameter of the power law equation $P(v)=C/v^\alpha $; b) a $p$-value rejecting or not the hypothesis $H_0$ that indeed our data could have been generated from a power law distribution.

@@

The table~\ref{tab:mualpha} summarizes the results obtained on all setups. Each value shown in the table is the mean values of 100 simulations. We see that in almost every case the null hypothesis cannot be rejected, which means that indeed the repartition of the price follows a power law. The only exception is for $mu$ equal to $0.064$, where the $p$-value is less than $0.05$. In this last case the null hypothesis is rejected which means that the distribution is not a power-law.

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[!h]
	\centering
	\begin{tabular}{ll|ccc}
		\multicolumn{2}{r}{Agents Number}&\multicolumn{2}{c}{Our result}&\multicolumn{1}{c}{Bentley et. al. 2004}\\
			&$\mu$ & $\alpha$ (sd) & $p$-value (sd) &$\alpha$ (sd)\\\hline
		N = 250	&0.004&1.53 (0.03)&0.8 (0.05)&1.54 (0.02)\\
			&0.016&1.57 (0.02)&0.6 (0.05)&1.57 (0.01)\\
			&0.064&1.66 (0.01)&0.0 (0.00)&1.67 (0.01)\\\hline
		N= 500	&0.004&1.50 (0.02)&0.6 (0.05)&1.53 (0.03)\\
			&0.016&1.55 (0.03)&0.3 (0.05)&1.61 (0.04)\\
			&0.064&1.78 (0.08)&0.0 (0.00)&1.81 (0.10)\\
	\end{tabular}
	\caption{Mean $\alpha$ \& SD are calculate on 100 runs for our results, and 5 runs for Bentley et. al. 2004.}
	\label{tab:mualpha}
\end{table}

For comparison purposes the results obtained by~\cite{bentley_random_2004} (which tested the ``neutral hypothesis'' with the same methodology) are added in the last column of table~\ref{tab:mualpha}. It appears that our results are highly similar to those obtained by~\cite{bentley_random_2004}. We note also that these results match the ones shown in~\cite{mesoudi_random_2009} where only one value of $\mu$ was tested ($mu = 0.008$). However, it is difficult to know if the slight differences observed between our work and those previous studies are statistically significant as the two previous studies rely on only five runs for the computation of the mean of $\alpha$ (against 100 in our case).

Nonetheless, for high value of $\mu$, previous works report that the distribution of variant follow a power law~\cite{bentley_random_2004}. This claim is justified based on the fact that the estimated $\alpha$ (estimated using linear regression on the log-log curve) has a high correlation coefficient. More recent works have shown that this method should be avoided~\cite{clauset2009powerlawdistributionsinempiricaldat}. Following these recent findings, our results preclude us to assume that the distribution of variant when $\mu$ is up to $0.064$ follow a power law .

A series of experiments are performed to see how the system reacts when multiple goods are present. We report in table~\ref{tab:multiGoods} the mean values of $\alpha$ and $p$-value obtained for 3 innovation rates (0.004, 0.016, 0.064) 4 number of goods (1, 3, 6, 9) and 250 agents.

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[h]
	\centering
	\begin{tabular}{l|llll}
		Innovation rate &\multicolumn{4}{c}{Number of goods}\\
		      & 1   & 3   &  6 & 9  \\\hline
		0.004 &1.53 (0.03)  &1.53 (0.03)&1.53 (0.03)&1.53 (0.03) \\
		0.016 &1.57 (0.02)  &1.57 (0.02)&1.56 (0.017)&1.57 (0.02) \\
		0.064 &1.66 (0.01)  &1.66 (0.01)&1.66 (0.01)&1.66 (0.01) \\\hline
	\end{tabular}
	\caption{Mean $\alpha$ and standard deviation for all innovation rate and different  of goods. The $\alpha$ given here is the mean $\alpha$ among all goods taken separately and computed over 100 runs with 250 agents.}
	\label{tab:multiGoods}
\end{table}

From the results shown in table~\ref{tab:multiGoods} it appears that for each innovation rate the distribution of variants has the same properties regardless of the number of goods. Additional analysis (not shown here) reveals that when the repartition of the price of every goods in a simulation with 9 goods is plotted, the 9 curves are superposed. Moreover these 9 curves are also superposed with the distribution shown in figure~\ref{fig:allMutation} for one good.

>>>>

We observe that the lower the mutation rate, the closer to a line the result is. This line corresponds to the ``power law" distribution explained in section~\ref{sec:culturalTrans}, and is typical of the result obtained under the ``neutral hypothesis". In order to verify if the distribution is in fact a power-law, we follow the method proposed by~\cite{clauset2009powerlawdistributionsinempiricaldat} and the R implementation proposed by~\cite{gillespie_fitting_2015}. Briefly, two values are returned: a) the estimation of the $\alpha$ parameter of the power law equation $P(v)=C/v^\alpha $; b) a $p$-value testing the null-hypothesis that our data could have been generated from a power law distribution.

@@

The table~\ref{tab:mualpha} summarizes the results obtained on all setups. Each value shown in the table is the mean values of 100 simulations. We see that in almost every case the null hypothesis cannot be rejected, which means that indeed the repartition of the price follows a power law. The only exception is for $mu$ equal to $0.064$, where the $p$-value is less than $0.05$. In this last case the null hypothesis is rejected, and we therefore assume that the distribution does not follow a power law.

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[!h]
	\centering
	\begin{tabular}{cc|ccc}
		\multicolumn{2}{r}{}&\multicolumn{2}{c}{Our result}&\multicolumn{1}{c}{Bentley et. al. 2004}\\
			N&$\mu$ & $\alpha$ (sd) & $p$-value (sd) &$\alpha$ (sd)\\\hline
		250	&0.004&1.53 (0.03)&0.8 (0.05)&1.54 (0.02)\\
			&0.016&1.57 (0.02)&0.6 (0.05)&1.57 (0.01)\\
			&0.064&1.66 (0.01)&0.0 (0.00)&1.67 (0.01)\\\hline
		500	&0.004&1.50 (0.02)&0.6 (0.05)&1.53 (0.03)\\
			&0.016&1.55 (0.03)&0.3 (0.05)&1.61 (0.04)\\
			&0.064&1.78 (0.08)&0.0 (0.00)&1.81 (0.10)\\
	\end{tabular}
	\caption{Mean $\alpha$ \& SD are calculate on 100 runs for our results, and 5 runs for Bentley et. al. 2004.}
	\label{tab:mualpha}
\end{table}

For comparison purposes the results obtained by~\cite{bentley_random_2004} (which tested the ``neutral hypothesis'' with the same methodology) are added in the last column of table~\ref{tab:mualpha}. It appears that these results are highly similar to ours. We note also that our results match the ones presented in~\cite{mesoudi_random_2009} where only one value of $\mu$ was tested ($mu = 0.008$). However, it is difficult to know if the slight differences observed between our work and those previous studies are statistically significant as the two previous studies rely on only five runs for the computation of the mean of $\alpha$ (against 100 in our case).

Nonetheless, for high value of $\mu$, previous works report that the distribution of variant follow a power law~\cite{bentley_random_2004}. This claim is based on the fact that the estimated $\alpha$ (estimation based on linear regression on the log-log curve) has a high correlation coefficient. Recent works have shown that the use of correlation coefficient should be avoided~\cite{clauset2009powerlawdistributionsinempiricaldat}. Following these recent findings, our results preclude us to assume that the distribution of variant when $\mu$ is up to $0.064$ follow a power law .

An additional series of experiments is done to see how the system reacts when multiple goods are present. We report in table~\ref{tab:multiGoods} the mean values of $\alpha$ and $p$-value obtained for 3 innovation rates (0.004, 0.016, 0.064), 4 number of goods (1, 3, 6, 9), and 250 agents.

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[!h]
	\centering
	\begin{tabular}{l|llll}
		Innovation rate &\multicolumn{4}{c}{Number of goods}\\
		      & 1   & 3   &  6 & 9  \\\hline
		0.004 &1.53 (0.03)  &1.53 (0.03)&1.53 (0.03)&1.53 (0.03) \\
		0.016 &1.57 (0.02)  &1.57 (0.02)&1.56 (0.017)&1.57 (0.02) \\
		0.064 &1.66 (0.01)  &1.66 (0.01)&1.66 (0.01)&1.66 (0.01) \\\hline
	\end{tabular}
	\caption{Mean $\alpha$ and standard deviation for all innovation rate and different  of goods. The $\alpha$ given here is the mean $\alpha$ among all goods taken separately and computed over 100 runs with 250 agents.}
	\label{tab:multiGoods}
\end{table}

From the results shown in table~\ref{tab:multiGoods}, it appears that for each innovation rate, the properties of the distribution of prices does not depend of the number of goods. Additional analysis (not shown here) reveals that when the distribution of prices of every goods in a simulation with 9 goods is plotted, the 9 curves are superposed. Moreover these 9 curves are also superposed with the distribution shown in figure~\ref{fig:allMutation} for one good.
